Errata

### 2025-04-06 tex/Manuscript-4.tex
Refactor code structure for improved readability and maintainability

>>>>

\begin{center}
    \textbf{\Large Manuscript 4}

    \vspace{0.3cm}
    \textit{Geoderma}\\
    \textit{DOI 10.2139/ssrn.4740477}

    \vspace{0.5cm}

    Julia Mitzscherling\textsuperscript{1*}, Rómulo Oses\textsuperscript{2}, Nicolás Riveras-Muñoz\textsuperscript{3}, Carsten W. Müller\textsuperscript{4,5}, Oscar Seguel\textsuperscript{6},   Peter Kühn\textsuperscript{3}, Thomas Scholten\textsuperscript{3} and Dirk Wagner\textsuperscript{1,7}

    \vspace{0.2cm}
  \end{center}

  \begin{small}
    \begin{justify}

        \textsuperscript{1}GFZ German Research Centre for Geosciences, Section Geomicrobiology, Telegrafenberg D-14473 Potsdam, Germany

        \textsuperscript{2}Centro Regional de Investigacion y Desarrollo Sustentable de Atacama, Universidad de Atacama (CRIDESAT UDA), Copayapu 484, Copiapó, Chile

        \textsuperscript{3}University of Tübingen, Soil Science and Geomorphology, Rümelinstrasse 19-23, D-72070 Tübingen, Germany

        \textsuperscript{4}Institute of Ecology, Chair of Soil Science, Technische Universität Berlin, 10587 Berlin, Germany

        \textsuperscript{5}Department of Geosciences and Natural Resource Management, University of Copenhagen, 1350 Copenhagen K, Denmark

        \textsuperscript{6}Universidad de Chile, Facultad de Ciencias Agronómicas, Av. Santa Rosa \#11315, La Pintana, 8820808 Santiago, 15 Chile

        \textsuperscript{7}University of Potsdam, Institute of Geosciences, Karl-Liebknecht-Strasse 24-25, D-14476 Potsdam, Germany

    \end{justify}
  \end{small}

  \vspace{0.5cm}
  \begin{center}
    Received: 27 February 2024
  \end{center}
  \cleardoublepage

\section*{Abstract}

While it is widely recognized that soil structure stability is related to soil physical, chemical and biological processes and properties, the role of the types of microorganisms in this respect is still unclear. Particularly, knowledge on the relative roles of abiotic versus biotic mechanisms is scarce. Our study investigated the influence of microbial activity on soil aggregate development across diverse climatic settings under wetting-drying (W-D) cycles and constantly moist conditions. Further, we evaluated the response of prokaryotic communities and microbial abundance to changing soil moisture conditions. Analyses encompassed soils from arid, semi-arid, mediterranean, and humid climates. Findings reveal the pronounced impact of climate legacy on soil structure responses to climate variations. Variations in soil aggregate attributes and microbial communities, driven by pedogenesis and climatic conditions, dictate how soil aggregation and microbial communities respond to changing soil moisture levels. In arid soils, the Birch-effect stimulated microbial activity and led to the breakdown of larger aggregates via labile organic matter consumption. Semi-arid soils showed microbial-mediated microaggregate formation through organic matter decomposition, ultimately decreasing soil stability. In mediterranean soils, increased microbial activity facilitated microaggregate assembly from silt-clay fractions, enhancing soil stability despite reductions in larger aggregates. Humid soils displayed unique responses, with fungi significantly enhancing stability, and prokaryotic activity stimulated by W-D. Conversely, microbial abundance substantially declined under constant moisture due to moisture stress, albeit a noteworthy increase in soil stability. Understanding these microbial-soil interactions as affected by climate is crucial for understanding initial soil formation processes as well as sustainable soil management, offering insights into soil health and ecosystem resilience.

\section*{Keywords} % Use \section* for unnumbered sections if needed
Soil aggregation, soil formation, Illumina MiSeq, microbial key players, microbial communities, wetting-drying cycles, climate gradient

\section{Introduction}

Most of the Earth's surface comprises soils and sediments, profoundly shaping the landscape by resisting erosion, influenced by the action of water in various climate regimes. Research exploring the interplay between biotic and abiotic processes in ecosystems has investigated organism-mineral interactions \citep{Brehm2005, Quirk2012} and biotic activities inducing soil and landscape formation \citep{Dietrich2006}. Biota impact terrestrial surfaces through mineral weathering, ion formation, and biofilm generation, which may control erosion resistance to a large extent \citep{Gadd2010, Gorbushina2007, RiverasMunoz2022}. Although recognized early that soil stability relates to physical, chemical, and biological processes, the specific role of microorganisms remains unclear, lacking a comprehensive understanding of their mechanisms and species contributions in varied climate contexts.

Typically, soils exhibit a complex three-dimensional arrangement comprising aggregates interspersed with void spaces \citep{Bailey2013, Oades1991}, increasingly referred to as soil architecture \citep{Amelung2023}. Aggregates consist of clusters of mineral particles and organic carbon, wherein the internal forces binding the particles together within each aggregate are much stronger than the forces acting between neighboring aggregates \citep{Six2004}. This strength enables the structures to endure wetting events and mechanical disruptions within the bulk soil \citep{Bravo2018}. These aggregates assemble hierarchically forming networks of particles and voids that are periodically connected during wetting episodes, which in turn create a variable flow of water and nutrients, offering access to soil organisms \citep{Wilpiszeski2019}.

In general, soil aggregates are grouped according to their size. Microaggregates (\SI{< 250}{\micro\metre}) consist of clay particles, organo-mineral complexes, and polyvalent metals, bound by bacterial polysaccharides \citep{Six2004, Totsche2018}, while macroaggregates (\SI{> 250}{\micro\metre}) are formations of minerals and particulate organic matter (POM), and are mainly bound by fungal hyphae and plant roots \citep{Oades1984}.

Soil aggregation is highly correlated with soil moisture, clay and organic carbon content \citep{AlKaisi2014} and is one controlling factor of the dynamic and spatial variation in soil properties \citep{Totsche2018}. The moisture regime in soils is largely controlled by climate which in turn controls soil formation. This has been shown in great detail along the Coastal Cordillera of Chile \citep{Bernhard2018b}. Climatic factors such as precipitation and temperature affect soil biological, soil chemical and soil physical processes such as the accumulation or organic carbon, weathering of minerals and swelling and shrinking of clay particles as well as soil properties. Terrestrial ecosystems experience regular changes in water availability, a critical resource for life. Alternating wet-dry events significantly influence soil aggregate formation and stabilization processes \citep{Najera2020}.

Additionally, microbial activity, i.e. the production of binding material such as exudates, and fungal hyphae affect aggregate formation \citep{Tisdall1982}, while soil aggregation also indirectly affects microbial activity as well as community structure \citep{Biesgen2020}. It has been shown that macro- and microaggregates are characterized by different predation pressure and, therefore, contain distinct microhabitats \citep{Fox2018, Mummey2004, Sessitsch2001}. Microbial residents occupy specialized niches within the aggregate structure, with active microorganisms living both within and between aggregate particles \citep{Bailey2013, Ebrahimi2016, Sessitsch2001}. The complex feedback between mineralogy and biology begins with the weathering of the parent material. It continues with the formation of aggregates and the stabilization of the microscale architecture of the soil structure and generally controls the geochemical cycles in the soil matrix during the development of a soil \citep{Crawford2012, Lynch1985}.

The influence of microorganisms on soil aggregate formation has been highlighted in several studies before \citep{Chotte2005, Gupta2015, Tisdall1982}. However, taxonomic groups driving the process and acting as potential key players are still largely unknown, while the mechanisms controlling the formation and persistence of soil aggregates are yet not fully understood. Particularly, knowledge of the relative roles of abiotic (water, organic-mineral-interactions) versus biotic (plant C input, microbial activity) drivers of aggregate formation is lacking. Therefore, we hypothesize that the abundance, diversity and composition of soil microbial communities are characteristic of different stages of soil development and as a corollary, that specific taxonomic groups are related to the distribution and stability of aggregates at each developmental stage.

Hence, this work aimed to (i) unveil the microbial contributions to soil aggregate formation and stabilization compared to abiotic processes in soils of different climatic conditions and developmental stages during repeated wetting-drying (WD) cycles, and (ii) identify microbial key players involved in aggregate formation and stabilization.

Since studies on the development of abiotic and biotic interactions are very complex, they require both different spatial and temporal scales \citep{Ollivier2011}. Therefore, soils were collected from four locations along the Coastal Cordillera of Chile representing a natural climate gradient, ranging from arid over semi-arid and mediterranean to humid conditions. In order to unveil the microbial role in the process of aggregate formation and stabilization, the manipulation of soil moisture by applying W-D cycles was performed with a set of native soils including the indigenous microbial communities and a set of sterilized soil samples without any microorganisms. Quantification of bacteria, archaea, and fungi by qPCR was used to document the response patterns of microbial abundances and communities. We performed Illumina MiSeq sequencing of 16S rRNA genes to gain a detailed picture of the microbial community structure and their shifts during W-D cycling. Changes in the microbial communities were related to edaphic parameters such as soil aggregate sizes, carbon and nitrogen content, C/N ratio and aggregate stability.

\section{Materials and methods}
\subsection{Field site description and sampling}

In the frame of the EarthShape project (German Science Foundation SPP 1803), soils were studied at the following four study sites distributed over \SI{1300}{\kilo\metre} along the Chilean Coastal Cordillera: Pan de Az{\'u}car National Park (\(\sim\)\ang{26}S), Santa Gracia Nature Reserve (\(\sim\)\ang{30}S), La Campana National Park (\(\sim\)\ang{33}S), and Nahuelbuta National Park (\(\sim\)\ang{38}S). 
The study sites cover a climate gradient with decreasing annual temperature and increasing annual precipitation from north to south and corresponded to arid, semi-arid, mediterranean and humid climates (\cref{fig:figure_1_label}) resulting in different soil moisture regimes \citep{Bernhard2018b}. 
Due to the long-term impact of stable climate conditions \citep{Ewing2006} soils show different stages of soil development. 
They were classified into Regosols in Pan de Az{\'u}car, Cambisols in Santa Gracia and La Campana, and Umbrisols in Nahuelbuta \citep{Bernhard2018b}. 
Vegetation types range from very sparse higher plant desert vegetation in the north, where biocrust soil coverage can reach as much as \SI{40}{\percent} of the area (biocrust-dominated desert vegetation) over semi-desert open shrubland and mediterranean dry forest and to temperate mixed broadleaved-coniferous forest in the south \citep{Bernhard2018a, Bernhard2018b}.

Surface depth increments (\SIrange{0}{5}{\centi\metre}) were collected in March--April 2016 on the mid slope position of the South-facing hillslopes in each study site and were directly sieved (\SI{< 2}{\milli\metre}) on ethanol-sterilized sieves to remove coarse organic material and stones. 
Samples were stored at \SI{4}{\degreeCelsius} until further processing. 
Detailed site-specific descriptions and soil classifications were performed by \citet{Bernhard2018a}, \citet{Bernhard2018b} and \citet{Oeser2018a}, \citet{Oeser2018b}.

\begin{figure}[H]
	\centering
	\includegraphics[width=1\textwidth]{img/M4-Figure_1.png}
	\caption{Location of the four study sites along the climate gradient on the Chilean Coastal Cordillera. MAP, mean annual precipitation; MAT, mean annual temperature.}
	\label{fig:M4-F1}
\end{figure}

\subsection{Experimental design}

Aggregate formation was developed during repeated W-D cycles using T{\"u}bingen cups (T-cup), which were covered by a fine silk cloth at the bottom \citep{Scholten2011}. 
In order to characterize the role of microorganisms on soil-structure formation processes, sets of soil samples with the indigenous microbial community (native) and sterilized control samples were prepared. 
The control samples were sterilized by cobalt-60 gamma-irradiation (SteriXpert, BBF Sterilisationsservice GmbH, Kernen-Rommelshausen, Germany). 
Each T-cup (\(h = \SI{5}{\centi\metre}\), \(r = \SI{2.5}{\centi\metre}\)) was filled with \SI{60}{\gram} soil material, followed by 22 days at \SI{22}{\degreeCelsius} to acclimate the soil-dwelling microorganisms to laboratory conditions. 
The manipulation experiment was performed by incubating the T-cups under controlled environmental conditions for a total time duration of six weeks. 
We saturated the soil to field capacity (pF 1.8, \SI{-60}{\hecto\pascal} pressure) at room temperature on a sterile sand bed (grain size \SIrange{0.06}{0.25}{\milli\metre}) for 24 hours. 
This was followed by a drying step at \SI{25}{\degreeCelsius} in an oven for 1--2 days until equilibrium. 
In addition, one sample remained on the sand bed throughout the entire experiment, being constantly subjected to moisture. 
Duplicate samples of native and sterilized soil material were collected after fulfilled cycles, comprising time point zero (R0), one, three, and six W-D cycles (R1, R3, R6), and after six weeks of constant moisture (CM). 
Separate T-cups were prepared for the different analytical approaches, one for DNA-based methods and one for the analysis of soil properties. 
Cups for DNA-analysis were stored at \SI{-20}{\degreeCelsius} and cups for the analysis of soil properties at \SI{4}{\degreeCelsius} until further processing.

\subsection{Soil properties, aggregate size distribution and stability}

Soil-chemical properties were analyzed for the native and the sterilized soils. 
The total organic carbon (TOC) was determined by Potsdamer Wasser- und Umweltlabor GmbH (PWU, Potsdam, Germany) according to DIN EN ISO/IEC 17025:2005. 
The aggregate size separation was done by wet sieving adapted from \citet{Elliott1986} as described in \citet{RiverasMunoz2022}. 
In the first step, \SI{10}{\gram} of soil material was irrigated with \SI{50}{\milli\liter} water. 
A series of four sieves, which were placed in descending order, was used to obtain the following soil aggregate size classes: 
(i) \SI{> 500}{\micro\metre} (large macroaggregates); 
(ii) \SIrange{250}{500}{\micro\metre} (small macroaggregates); 
(iii) \SIrange{53}{250}{\micro\metre} (large microaggregates); 
(iv) \SIrange{20}{53}{\micro\metre} (small microaggregates), and 
(v) \SI{< 20}{\micro\metre} (silt/clay). 
Aggregates were separated manually by moving the sieve up and down 30 times during \SI{5}{\minute} in a \SI{2.5}{\liter} water tank. 
The aggregate classes retained on each sieve were oven-dried at \SI{105}{\degreeCelsius}, and weighed. 
In the following, large and small macroaggregates are referred to as macroaggregates (MAC), large and small microaggregates are referred to as microaggregates (MIC) and the silt/clay class is referred to as SC.

The mean weight diameter (MWD in \si{\milli\metre}), a measure of aggregate stability determined by the size distribution of aggregates \citep{Amezketa1999}, was calculated as an average weighed by aggregate size.

Furthermore, the soil organic carbon (SOC) and total nitrogen (TN) content of each aggregate size class was measured using oxidative heat combustion at \SI{1150}{\degreeCelsius} in a Vario EL III elemental analyzer (Elementar Analysensysteme GmbH, Hanau, Germany) and subsequently, the C/N ratio was calculated. 
SOC and total carbon were used interchangeably since the study sites showed low inorganic carbon content \citep{RiverasMunoz2022}.

\subsection{DNA extraction and quantification of bacteria, archaea, and fungi}

Total genomic DNA was extracted from the native soil material using the DNeasy\textsuperscript{\textregistered} PowerSoil\textsuperscript{\textregistered} Kit (Qiagen GmbH, Hilden, Germany) with a maximum amount of \SI{0.25}{\gram} per sample. 
DNA was extracted in triplicates following the manufacturer’s protocol with one exception: DNA elution was done with PCR-grade water. 
Final DNA concentrations were measured with a Qubit\textsuperscript{\textregistered} 2.0 Fluorometer using the Qubit\textsuperscript{TM} dsDNA HS Assay Kit (both Invitrogen Life Technologies, Thermo Fisher Scientific, Carlsbad CA, USA). 
DNA extracts were stored at \SI{-20}{\degreeCelsius} until further processing.

Gene copy numbers of bacteria, archaea, and fungi were determined by quantitative polymerase chain reaction (qPCR). 
16S rRNA genes were amplified by using the bacterial primer pair Eub341F (5\('\)-\texttt{CCT ACG GGA GGC AGC AG}-3\('\)) and Eub534R (5\('\)-\texttt{ATT ACC G CG GCT GCT GG}-3\('\)) \citep{Muyzer1993} and the archaeal primer pair 340F (5\('\)-\texttt{CCC TAC GGG GYG CAS CAG}-3\('\)) and 1000R (5\('\)-\texttt{GGC CAT GCA CYW CYT CTC}-3\('\)) \citep{Gantner2011}. 
28S rRNA genes were amplified by using the fungal primer pair NL1F (5\('\)-\texttt{ATA TCA AAT AAG CGG AGG AAA AG}-3\('\)) and LS2R (5\('\)-\texttt{ATT CCC AAA CAA CTC GAC TC}-3\('\)) \citep{Bates2009}. 
The qPCR assay was carried out in a total reaction volume of \SI{20}{\micro\liter} using the KAPA SYBR\textsuperscript{\textregistered} FAST qPCR Kit Master Mix (2x) Universal (Kapa Biosystems, Sigma-Aldrich, Germany). 
Quantification was performed in the CFX96 Connect\textsuperscript{TM} Real-Time System (Bio-Rad Laboratories, CA, USA) with the following cycling program for bacteria/ archaea/ fungi: initial denaturation at \SI{95}{\degreeCelsius} for \SI{3}{\minute}, followed by 40/ 45/ 45 cycles of denaturation at \SI{95}{\degreeCelsius} for \SI{3}{\second}, annealing at \SI{60}{\degreeCelsius}/ \SI{57}{\degreeCelsius}/ \SI{50}{\degreeCelsius} for \SI{20}{\second}, and elongation at \SI{72}{\degreeCelsius} for \SI{30}{\second}. 
Fluorescence was measured at \SI{80}{\degreeCelsius}. 
The melting curve was recorded by rising temperature from \SI{65}{\degreeCelsius} to \SI{95}{\degreeCelsius}. 
DNA extracts from Pan de Az{\'u}car were added undiluted, samples from La Campana and Santa Gracia were used in 1:100 dilutions, as well as samples from Nahuelbuta with one exception: constant moisture samples were used in 1:10 dilutions. 
Every qPCR run included blanks, calibration standards, and soil samples in triplicates. 
The reaction efficiency ranged between \SIrange{96.3}{103.9}{\percent}. 
Data were analyzed using the CFX Manager\textsuperscript{TM} Software (Bio-Rad Laboratories, CA, USA).

\subsection{Illumina MiSeq amplicon sequencing of bacteria and archaea}

PCR products were prepared for Illumina MiSeq high-throughput sequencing of bacterial and archaeal 16S rRNA genes. 
Unique primer combinations of barcode tagged 515F (5\('\)-\texttt{GTG CCA GCM GCC GCG GTA A}-3\('\)) and 806R (5\('\)-\texttt{GGA CTA CGV GGG TWT CTA AT}-3\('\)) primers were assigned to each sample \citep{Caporaso2012}. 
For each genomic DNA sample, two technical replicates with different barcodes were amplified by PCR and sequenced. 
For a detailed description of the library preparation, the reader is referred to \citet{Moskwa2020}.

\subsection{Bioinformatics and statistical analysis}

Sequencing was performed at Eurofins Genomics (Konstanz, Germany) on an Illumina MiSeq generating \(2\times 300\)\,bp paired-end reads. 
The sequencing library was demultiplexed using \texttt{cutadapt} v3.4 \citep{Martin2011} using the following parameters: \texttt{-e 0.2 -q 15,15 -m 150 --discard-untrimmed}. 
The ASVs were generated using trimmed reads and the \texttt{DADA2} package v1.20 \citep{Callahan2016} with R v4.1 using the pooled approach with the following parameters: \texttt{truncLen=c(240,200), maxN=0, rm.phix=TRUE, minLen=200}. 
Taxonomic assignment was done using \texttt{DADA2} and the SILVA database v138 \citep{Quast2012}. 
Subsequently, ASVs representing chloroplasts, mitochondria and singletons were removed.

Prior to statistical analysis, absolute read counts were transformed into relative abundances. 
The similarity of duplicate samples was evaluated visually in a non-metric multidimensional scaling (NMDS) plot before duplicates were merged for microbial community analysis. 
The microbial community structure and diversity as well as its relation to the soil properties was assessed using the Past 4.09 software \citep{Hammer2001}. 
Beta diversity of the microbial communities and possible associations between community structure and environmental variables were assessed in a Canonical correspondence analysis (CCA) ordination plot using the Bray-Curtis dissimilarity. 
Soil edaphic properties were normalized using z-score transformation. 
Principal component analyses (PCA) based on Euclidean distance were used to assess the variation of edaphic parameters across the four study sites and to test for variation between native and sterile soil material of each site. 
The significance of variation between native and sterile soil material was tested using one-way PerMANOVA. 
Pearson and Spearman correlation were used to evaluate associations of the abundance of microbial taxa on different taxonomic levels with the edaphic parameters. 
Differences of soil aggregate, SOC and TN content as well C/N ratio between native and sterile soil material were assessed using ANOVA. 
Differences in the family level community composition were tested using the Kruskal-Wallis and Mann-Whitney post-hoc tests. 
The FAPROTAX database \citep{Louca2016} was used to predict functions from the ASV-level taxonomy.

\subsection{Data deposition}

Sequencing raw data are publicly available via the European Nucleotide Archive under the Project accession number \texttt{PRJEB72146}: \url{https://www.ebi.ac.uk/ena/browser/view/PRJEB72146}. 
All data used in this study are freely available under the Creative Commons Attribution 4.0 International (CC BY 4.0) open access license at GFZ data services. 
All samples are identified via a unique IGSN (international geo sample number). 
These data are available as a supplementary dataset in \citet{Bernhard2018b} with the following DOI: \url{https://doi.org/10.5880/GFZ.5.3.2018.001}. 
When using the data please cite: \citet{Bernhard2018a}.

\section{Results}
\subsection{Soil properties}

\begin{figure}[H]
	\centering
	\includegraphics[width=1\textwidth]{img/M4-Figure_2.png}
	\caption{Soil properties in native soils (+) and sterilized soils (-) of different climates. 
    A) Aggregate size distribution. 
    B) SOC content, 
    C) TN content, and 
    D) C/N ratio per aggregate class. 
    MAC, macroaggregate fraction \SIrange{250}{2000}{\micro\metre}; 
    MIC, microaggregate fraction \SIrange{20}{250}{\micro\metre}; 
    SC, silt/clay fraction \SI{< 20}{\micro\metre}. 
    Significant differences between native and sterilized soils are indicated by asterisks (* \(p < 0.05\); ** \(p < 0.01\)).}
	\label{fig:M4-F2}
\end{figure}

Macroaggregates (MAC), microaggregates (MIC) and the silt/clay (SC) classes were differently distributed in arid, semi-arid, mediterranean and humid soils (\cref{fig:figure_2a_label}). 
Less developed soils from arid and semi-arid climate conditions were characterized by the lowest content of MAC ranging around \SIrange{60}{65}{\percent}. 
Semi-arid soils displayed a higher MIC content and a lower SC content compared to arid soils. 
More developed soils from the mediterranean and humid climates were characterized by the highest abundance of MAC with nearly \SI{80}{\percent} in mediterranean soils and more than \SI{90}{\percent} in humid soils. 
Accordingly, the proportion of MIC was low in mediterranean soils (\(\sim\)\SI{20}{\percent}) and lowest in humid soils (\SI{< 20}{\percent}). 
Moreover, SC was hardly present in mediterranean and humid soils (\(\sim\)\SI{2}{\percent}).

In native soils, including the indigenous microbial communities MAC contents decreased in arid, semi-arid and mediterranean soils, but increased in humid soils. 
MIC contents remained stable in arid soils, increased in semi-arid and mediterranean soils, but decreased in humid soils. 
The SC class slightly increased in arid soils, remained stable in semi-arid and mediterranean soils, and decreased in humid soils.

The SOC and TN (\cref{fig:figure_2b-c_label}) contents were lowest in the less developed soils and increased with soil development along the climate gradient, with the lowest SOC (\(\sim\)\SI{1}{\percent}) and TN values (\(\sim\)\SI{0.1}{\percent}) in arid soils and highest values in humid soils (\(\sim\)\SI{10}{\percent} and \(\sim\)\SI{0.5}{\percent}). 
In each soil, the SOC and TN contents were lowest within the MAC class and highest within the SC class. 
The C/N ratios (\cref{fig:figure_2d_label}) were highest in arid soils. 
Lowest values were observed in semi-arid soils with an increasing trend over mediterranean to humid soils. 
Within each soil, the highest C/N ratios can be observed within MAC, followed by SC and MIC.

The aggregate stability (MWD) was lowest in arid and semi-arid soils, increased in mediterranean soils and peaked in humid soils (\cref{fig:figure_a1_label}).

A principal component analysis (PCA) including soil aggregate size classes, aggregate stability (MWD), SOC, TOC and TN content as well as the C/N determined during repeated W-D cycles (\cref{fig:figure_3a_label}) revealed similar soil properties in native and sterilized soils from the same climatic condition. 
The separation of soils along the x-axis, explaining \SI{66.0}{\percent} of the variance, was in accordance with the climate gradient from arid to humid conditions. 
The difference between arid and semi-arid soils was mainly visible on the y-axis, explaining \SI{27.6}{\percent} of the variance. 
In addition, native and sterilized soils from arid conditions were separated along the y-axis.

\begin{figure}[H]
	\centering
	\includegraphics[width=1\textwidth]{img/M4-Figure_3.png}
	\caption{Soil properties and microbial community structure in soils of different climates. A) Principal Component Analysis of soil properties in native soils (squares) and sterilized soils (circles) including MWD, TOC, soil aggregates, SOC, TN and C/N per aggregate size class. B) Canonical correspondence analysis of the ASV composition including soil properties: MWD, aggregate class (MAC, MIC, SC) and the respective C/N ratio, SOC and TN content. R0-R6; number of W-D repetitions; CM, constant moisture. Colors represent different climatic conditions.}
	\label{fig:M4-F3}
\end{figure}

\subsection{Microbial influence on soil properties}

Differences between native and sterilized soils were further evaluated in separate PCA ordinations, showing significant differences between native soils and sterile controls not only in arid (\(p = 0.005\)), but also in semi-arid (\(p = 0.008\)) and mediterranean soils (\(p = 0.02\)) (\cref{fig:figure_4_label}, \cref{tab:table_a1_label}). 
At these three sites, the native samples with microorganisms before W-D treatment (\(\square\)R0) were similar to all sterile controls (\(\circ\)R0--R6, CM) and were thus referred to as microbially not (yet) influenced. 
After at least 1 W-D cycle, microbial samples clustered apart from the sterile controls and were thus claimed to be microbially influenced. 
The difference between microbially influenced and not influenced samples appeared mainly along the x-axis of the plots, which explained the highest variance. 
In arid soils, the microbially influenced soils were characterized by higher C/N ratios in all aggregate classes (\cref{fig:figure_4a_label}). 
In semi-arid and mediterranean soils, soil aggregate sizes in microbially influenced soils were characterized by higher levels of MIC and C/N\textsubscript{SC}, and lower levels of MAC compared to the sterile controls (\cref{fig:figure_4b-c_label}). 
Microbially influenced humid soils were not significantly different from sterile soils. 
However, the majority of microbially influenced soils was characterized by higher MAC and MWD, while sterile soils had higher MIC contents (\cref{fig:figure_4d_label}).

\begin{figure}[H]
	\centering
	\includegraphics[width=1\textwidth]{img/M4-Figure_4.jpg}
	\caption{Principal Component Analyses of soil properties in native soils (squares) and sterilized soils (circles) including significant soil properties. A) arid, B) semi-arid, C) mediterranean and D) humid site. Microbial influence is indicated by sample separation along the x-axis.}
	\label{fig:M4-F4}
\end{figure}

\subsection{Microbial abundance}

Based on 16S and 28S rRNA gene copy numbers, bacteria were most abundant among all four climate conditions, followed by archaea and fungi (\cref{fig:figure_5a_label}). 
Only in humid soils, fungi were slightly more abundant than archaea and possessed highest abundances among soils. 
The lowest abundances of bacteria, archaea and fungi were observed in arid soils. 
Still, they increased over the course of the experiment with a peak after 3 cycles and maximum values under constant moisture. 
Semi-arid and mediterranean soils were characterized by highest abundances that remained relatively constant during the W-D cycles and constant moisture. 
In humid soils, bacterial and archaeal abundances were stable and fungal gene copies increased after 3 W-D cycles. 
However, all gene abundances of humid soils were lowest in soils exposed to constant moisture.

Linear (Pearson) and non-linear (Spearman) correlation analyses (\cref{fig:figure_a2_label}) revealed that bacterial and archaeal abundances were positively correlated with each other. 
Both were inversely correlated with the C/N ratios in each soil class. 
On the contrary, the fungal abundance was positively correlated with MAC and MWD, while being inversely correlated with MIC and the S/C class. 
Furthermore, fungi were positively correlated with the SOC and TN contents in each aggregate class.

\begin{figure}[H]
	\centering
	\includegraphics[width=1\textwidth]{img/M4-Figure_5.jpg}
	\caption{Prokaryotic and fungal abundance as well as prokaryotic diversity in soil material from different climates. A) 16S and 28S rRNA gene abundances of bacteria, archaea and fungi. Bacterial and archaeal diversity based on B) Shannon and C) Chao-1 diversity indices. R0 - R6, number of W-D repetitions; CM, constant moisture}
	\label{fig:M4-F5}
\end{figure}

\subsection{Bacterial and archaeal diversity}

Similar to their abundance, bacterial and archaeal diversity based on the alpha-diversity indices Chao-1 and Shannon (\(H\)) was highest in semi-arid soils from Santa Gracia (median \(H = 7.4\)) and in mediterranean soils of La Campana (\(H = 7.5\)) with maximum diversity indices in mediterranean soils (\cref{fig:figure_5b_label}). 
Although lowest microbial abundances were observed in arid soils, the bacterial and archaeal diversity was lowest in humid soils from Nahuelbuta (\(H = 6.0\)), followed by arid soils from Pan de Az{\'u}car (\(H = 6.6\)). 
In arid and semi-arid soils, the diversity according to Chao-1 increased after one W-D cycle and decreased afterwards. 
Contrary, mediterranean and humid soils were characterized by an increasing diversity during W-D cycles, but possessed a low diversity in the sample exposed to constant moisture. 
Like the bacterial and archaeal abundance, the number of taxa and the diversity were negatively correlated with the C/N ratios in all soil classes (\cref{fig:figure_a3_label}).

\subsection{Microbial community structure and composition}

Like the soil edaphic properties, the microbial communities clustered according to the climatic condition of each study site (\cref{fig:figure_3b_label}). 
In semi-arid and mediterranean soils, microbial communities did not significantly change during the treatment. 
Both communities were similar and clustered apart from those of arid and humid soils along CCA 2. 
While microbial communities of the semi-arid and mediterranean soils were similar (along CCA 1), soil parameters of the respective sites differed significantly (along PC 1). 
In contrast, microbial communities in humid soils displayed an increasing variance along the y-axis (explaining \SI{24.3}{\percent} of the variance) with increasing number of W-D cycles and under constant moisture. 
Similarly, the microbial communities of R6 and CM in arid soils varied from R0--R3 along the x-axis (explaining \SI{26.5}{\percent}).

The microbial community composition was assessed by MiSeq amplicon sequencing. 
After quality check and filtering (including removing reads of the negative control) the dataset comprised 1.2 million reads across \(\sim\)\num{11000} amplicon sequencing variants (ASVs). 
From this \(\sim\)\SI{98.1}{\percent} of the reads were assigned to bacteria and \SI{1.9}{\percent} to archaea. 
After taxonomic classification, 828 putative genera (807 for bacteria, 21 for archaea) were obtained. 
Based on phylogenetic classification, ASVs could be assigned to the ten dominant prokaryotic phyla Proteobacteria (\SI{24.8}{\percent} of total ASVs), Actinobacteriota (\SI{22.9}{\percent}), Acidobacteriota (\SI{11.9}{\percent}), Chloroflexi (\SI{8.7}{\percent}), Planctomycetota (\SI{8.5}{\percent}), Bacteroidota (\SI{5.1}{\percent}), Gemmatimonadota (\SI{4.8}{\percent}), Verrucomicrobiota (\SI{3.6}{\percent}), Firmicutes (\SI{2.3}{\percent}) and Crenarchaeota (\SI{1.8}{\percent}) (\cref{fig:figure_a4_label}).

The bacterial communities of all four sites comprised 63 families and 18 unknown orders (\cref{fig:figure_6_label}, \cref{tab:table_a2_label}). 
Of these families, 32 possessed almost or more than \SI{2}{\percent} mean relative abundance per site. 
Some were abundant (\SI{> 2}{\percent}) only at one site, while others were found in soils from several climatic conditions. 
Euzebyaceae, Pseudonocardiaceae, Caldilineaceae, AKYG1722 and JG30-KF-CM45 (Thermomicrobiales), Trueperaceae, Longimicrobiaceae, Pirellulaceae and unclassified Rhizobiales were abundant exclusively in arid soils from Pan de Az{\'u}car. 
Nitrososphaeraceae, Geodermatophilaceae, Micromonosporaceae and Beijerinckiaceae were abundant in the semi-arid soils of Santa Gracia. 
Vicinamibacteraceae and unclassified Vicinamibacterales were found to be abundant in mediterranean soils of La Campana. 
In contrast, Subgroup 1 of Acidobacteriales, Acidothermaceae, Ktedonobacteraceae, Gemmataceae and unclassified Elsterales were abundant only in the humid soils of Nahuelbuta.

\begin{figure}[H]
	\centering
	\includegraphics[width=1\textwidth]{img/M4-Figure_6.png}
	\caption{Relative abundance of prokaryotic families in soil material from different climates and after different soil treatments. R0-R6; number of W-D repetitions; CM, constant moisture. Each bubble represents the mean relative abundance of technical replicates.}
	\label{fig:M4-F6}
\end{figure}

\subsection{Predicted microbial functions}
